%\documentclass[12pt,letterpaper]{article}
% JOURNAL STUFF
\documentclass[global]{svjour}
\journalname{Engineering with Computers}

%\usepackage{imr}
%\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{url}
\usepackage{float}
\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}
\floatname{algorithm}{Algorithm}

%\def\thepage {}
%\bibliographystyle{imr}
%\newtheorem{Pseudo}{Pseudo-Code}[section]
\begin{document}

\title{Toward Interoperable Mesh, Geometry and Field Components for PDE Simulation Development}

%% \author{Kyle K. Chand \and 
%%   Lori Freitag Diachin \and 
%%   Xiaolin Li \and
%%   Carl Ollivier-Gooch\and
%%   Mark Shephard \and
%%   Timothy Tautges \and
%%   Harold Trease}
% JOURNAL STUFF
\titlerunning{Interoperable Components for Simulation Development}
\author{Kyle K. Chand\inst{1} \and 
  Lori Freitag Diachin\inst{1} \and 
  Xiaolin Li\inst{2} \and
  Carl Ollivier-Gooch\inst{3}\and
  E. Seegyoung Seol \inst{4}\and
  Mark S. Shephard\inst{4} \and
  Timothy Tautges\inst{5} \and
  Harold Trease\inst{6}}
\institute{Center for Applied Scientific Computing, Lawrence Livermore National Laboratory \and
  Dept. of Applied Mathematics and Statistics, SUNY Stonybrook \and
  Advanced Numerical Simulation Laboratory, University of British Columbia \and
  Scientific Computation Research Center, Rensselaer Polytechnic Institute \and
  Mathematics and Computer Science Division, Argonne National Laboratory \and
  Pacific Northwest National Laboratory}

\maketitle

\begin{abstract}
Mesh-based PDE simulation codes are becoming increasingly
sophisticated and rely on advanced meshing and discretization tools.
Unfortunately, it is still difficult to interchange or interoperate
tools developed by different communities to experiment with various
technologies or to develop new capabilities.  To address these
difficulties, we have developed component interfaces designed to
support the information flow of mesh-based PDE simulations.  We
describe this information flow and discuss typical roles and services
provided by the geometry, mesh, and field components of the
simulation.  Based on this delineation for the roles of each
component, we give a high-level description of the abstract data model
and set of interfaces developed by the Department of Energy's
Interoperable Tools for Advanced Petascale Simulation (ITAPS)
center.  These common interfaces are critical to our interoperability
goal, and we give examples of several services based upon these
interfaces including mesh adaptation and mesh improvement.
\end{abstract}

\section{Introduction}
\input{intro}

\section{Information Flow in Mesh-based Simulation}\label{sec:infoFlow}
\input{simulation}

\section{The ITAPS Interface Definition efforts}\label{sec:tsttdef}
\input{interface}

\section{ITAPS interface use cases}\label{sec:apps}

The ITAPS data model and interfaces have been defined and
implementations are underway at many different institutions.
In particular, the mesh interface is the most mature and in this
section we give several examples of its use in adaptive loop
construction for two different applications and in mesh quality improvement
tools.  The appendix provides code for two elementary examples illustrating 
some simple uses of the mesh interface.

\input{adapt_loop}
\input{improve}

\section{Concluding Remarks}
A simulation's information flow provides a conceptual framework for
designing interoperable tools for geometry management, mesh generation
and discretization.  Using this framework, the Interoperable Tools for
Advanced Petascale Simulation center has developed a set of language
independent interfaces to geometry, mesh, and solution field
information.  Several groups have successfully implemented the ITAPS
mesh interface with a diverse range of technologies ranging from
structured composite grids to fully unstructured infrastructures.
These implementations have provided mesh services for ITAPS-based
interoperable components providing technologies such as mesh
adaptation, design optimization and mesh improvement.  In principle,
any implementation of the the ITAPS mesh interface now has access to
these advanced technologies without requiring new source code
development.

While the utility of these interfaces has been demonstrated in a
number of applications, the current work remains a proof of principle.
Advancing the interface definition to a reliable standard requires
further investigation and demonstration.  For example, the nascent
solution field and operator interfaces have yet to be completed and
implemented.  In addition, more implementations of the interfaces need
to be created and exercised to make extensive interoperability and
interchangeability a possibility.  Furthermore, the performance
ramifications of using these interfaces must be carefully examined in
order to assure that applications built upon this infrastructure are
not plagued by low performance.  Collaborations with the Common
Component Architecture group~\cite{cca-forum} will lead to the
development of higher level components that advanced services to
simulation code developers.

More information on the ITAPS center can be found at
http://www.itaps-scidac.org/.

%\section*{Acknowledgments} 
% JOURNAL STUFF
\begin{acknowledgement}

We would like to thank Tamara Dahlgren of LLNL for her insightful
comments on the ITAPS interface design.  

This work was performed under the auspices of the U.S. Department of
Energy by the University of California Lawrence Livermore National
Laboratory under contract No. W-7405-Eng-48 (UCRL-JRNL-213577); 
the Canadian Natural Sciences and Engineering Research 
Council under Special Research Opportunities Grant SRO-299160;
and by Rensselaer Polytechnic Institute under DOE grant 
number DE-FC02-01ER25460. 

%WE NEED auspices for SANDIA, STONYBROOK, RPI , and PNNL

% JOURNAL STUFF
\end{acknowledgement}

\bibliographystyle{plain}
\bibliography{../../biblio/tstt.bib}

\input{examples}
\end{document} 
