In mesh-based analysis procedures thePDEs are solved approximately through a double discretization processin which the problem's physical domain is discretized into a set ofpiecewise components (e.g. a mesh) and the PDEs to be solved arediscretized over the mesh in an appropriate manner. Thediscretizations of the PDEs over the mesh are assembled into a fullydiscrete system that is solved. The result of thisprocess yields the construction of a set of discretized solutionfields. Methods covered by such approaches include finitedifference, finite volume, finite element, boundary element andpartition of unity (so-called meshfree) methods.\subsection{Problem Definition}To qualify the operations and information involved in executing amesh-based simulation, we begin with a qualification of the problemdefinition which includes:\begin{itemize}\item The domain over which the simulation is to be solved. For theclasses of simulations being considered here the domain includes aspatial component that is one-, two-, or three-dimensional. The problemcan also be defined over time in which case the domain also includes atemporal component.\item The mathematical form governing the simulation (PDE’s,variational principle, weak form).\item Specification of the parameters, referred to as physicalattributes, associated with the governing mathematical equations thatincludes:\begin{itemize}\item material properties\item forcing functions\item boundary conditions\item initial conditions\end{itemize}\end{itemize}From the viewpoint of supporting a numerical simulation, thedomain representation must be able to:\begin{itemize}\item Support the construction of a mesh that represents the domainlevel discretization used in the simulation.\item Support the ability to address any geometry interrogation required.\item Support the proper association of the physical and mathematicalattributes with the mesh.\item Support the domain evolution in the cases where the domainchanges as part of the solution process.\end{itemize}There are multiple sources for the high level definitions of thespatial component of the domain with CAD models, image data andcell-based (mesh-based) being the most common. Each of these sourceshas one or more representational forms. Historically, CAD systems useboundary representations. Image data use a volumetricform such as voxels or octrees. Depending on the configuration of thecells a variety of implicit and explicit boundary or volumetricrepresentations have been used.Except in cases of image data and when all aspects of the simulationprocess can be effectively defined in terms of volume entities, itis generally accepted that the use of a boundaryrepresentation is well suited for the spatial domain definition. Thereis a substantial computer-aided design literature on the variousboundary representations. Common to these representations isthe use the abstraction of topological entities and their adjacenciesto represent the model entities of different dimensions. In aboundary representation the information defining the shape ofthe topological entities can be thought of as attribute informationassociated with the appropriate entities. The ability to interact withtopological entities provides aneffective means to develop abstract interfacesallowing the easy integration to multiple domain definitionsources.An important consideration in the selection of a boundaryrepresentation is its ability to represent the classes of domainneeded. In the case of numerical simulations the domains to be meshedcan be general combinations of 0-, 1-, 2- and 3-D entities in generalconfigurations. Figure 1 shows a typical analysis domain that may beused for the structural analysis of a portion of a pipingsystem. The analysis domain is an idealizationwhere portions of the pipes are idealized by beams,the support bracket idealized by a plateand a full 3-D solid used for the pipe juncture.Figure 1. Example of a non-manifold model used in simulation.The proper representation of such geometric domains, aswell as others like multi-material domains, arereferred to as a non-manifold boundary representations \cite{GuCh90,We88}. Inthe case of non-manifold models the representation must indicatehow topological entities are used by bounding higher orderentities. For example, each side of a face may be used by a differentregion. Therefore, faces have two uses. Another terminology for theuse of a topological entity by higher order topological entities isco-entities \cite{Ta00}.Geometric modeling systems maintain tolerance information on hownumerically well the entities fit together.This is necessitated by the fact that to function properly geometricmodeling systems must employ finite tolerances.The algorithms and methods within the geometricmodeling system use the tolerance information toeffectively define and maintain a consistent representation of thegeometric model. (What many geometry-basedapplications have referred to as dirty geometry is caused by a lack ofknowledge and proper use of the tolerance information \cite{BeWa04}.)The abstraction of topology provides an effective means to developfunctional interfaces to boundary-based modelers. The ability togeneralize these interfaces is further enhanced by the fact thatthe geometry shape information needed by most allsimulation procedures consists of pointwise interrogations that can be easily answered in a method independent of the modeler shape representation.The developer of CAD systems supportgeometry-based applications through general APIs.These geometric modeling APIs have been successfullyused to developed automate finite element modeling processes\cite{BeWa04,ShGe92}.In some cases the only domain representation is amesh. In these cases it is still desirableto construct a high level topological representation of the problemdomain. In this case theprocess of constructing, or updating, the topological entitiesassociated with the domain geometric model is focusedon determining the appropriate sets of mesh regions, faces, edges, andvertices to associate with the model regions, faces, edges andvertices respectively. Algorithms to do this based on mesh basedgeometry and/or simulation contact or fracture informationhave been developed \cite{KrOr01,PaOr02,WaSh05}.Once the model topology has been set,the geometric shape information can be defined in terms ofthe mesh facets, or can be made higher order \cite{CiOr00,OwWh01}.An examination of the properties of analysis attributes indicatesthey are tensorial quantities \cite{BeSo83} thatmust be defined with respect to a coordinate system. Generalized structures and methods can define analysis attributes and associate themgeometric model entities \cite{OBSh02}.\subsection{Domain Discretization}The mesh, is a piecewise decomposition of the space/time domain. The specifics of the definition of the mesh is a function of the methods used to discretize the equations over the mesh entities.It is common to employ different discretizations for the spatial andtemporal domains.  Since the definition of the spatial mesh is typically the more complex of the two, it is the focus ofthis discussion. The requirements of the mesh are:\begin{itemize}\item To have the appropriately defined “union” of the meshentities represent the domain of interest.\item To maintain, or have access to, the geometric shape informationneeded for processes such as differentiation and integration.\item To support the PDE discretization process over the mesh entities.\item To maintain relationships of the mesh entities needed to supportthe assembly of the complete discrete system and construction of thesolution fields.\end{itemize}One common mesh form is the conforming mesh where the intersections oftwo mesh entities is null and the intersections of their closure iseither null or the closure of a common boundary mesh entity (face,edge or vertex). A variant of this is the non-conforming mesh wherethe intersections of the closure of two mesh entities is null, ordifferent parts of the boundaries of the two mesh entities. Other meshstructures employ mesh patches that can interact in a variety ofways. Finally, other methods are defined in terms of overlappingregions (e.g., spheres or cubes). In each of these cases there arerules on how the mesh entities interact, howequation discretizations are performed over them, and how the completediscrete system is assembled.The geometric shape of the mesh entities must beunderstood to support the equation discretization process. Inmany methods the mesh geometry is implied from a discrete set ofparameters, which is satisfactory for fixed mesh simulations. Analternative is to maintaina linkage to the high level domain geometry. This alternative tends tobe expensive soa mesh based geometric definition is typically used. However,in the case of adaptive mesh improvements it is necessary to use thelinks back to the original domain geometry to ensure the meshgeometric approximation improves in a manner consisted with the orderof accuracy provided by the equation discretization process. Forexample, as piecewise linear elements approximating curved portions ofthe geometry are refined, the new mesh vertices need to be placed onthe curved boundary, or as the polynomial orderof an element is increased, the geometric approximation of the closureof that entity must be increased to the correct order.The data model for the mesh must maintain an association with thedomain definition, the discretization functions, the assembleddiscrete system and the solution fields. From the perspective ofmaintaining its relationship to the geometric domain, the useof topological entities and their adjacency is ideal\cite{BeSh97,DeOB01,Ta00}. In this manner it is possible to associatethe mesh entities to the domain entities to obtain needed attributesand geometric information. This association between the two modelstructures is referred to as classification below.In other cases, such a representation is not ideal.For example, something like an octree, or some otherspatially-based structure, is appropriate for the partition of unity(so call meshfree) methods. In the case of structured meshesmaintaining a complete topology down to the individual cell entitieswould be overkill.However, in both of these cases there is information that is wellsuited to a topological representation. For example, the boundariesof the mesh patches in an structured mesh are ideallydefined in terms of a topological structure augmented with the rulesof mesh patch interaction. In the case of partition of unitymethods, maintaining topological entities of the cells of anoctree effectively supports the needed operations \cite{KlSh00}.Consider the case of using a topological structure for the mesh.Under the assumption that each topological meshentity of dimension $d$, $M^d_i$, is bounded by a set of topologicalmesh entities of dimension $d-1$, $\left\{ M^d_i \left\{ M^{d-1} \right\} \right\}$, the full set of mesh topological entities are:\begin{equation}T_M = \left\{ \left\{ M \left\{ M^0 \right\} \right\},~\left\{ M \left\{ M^1 \right\} \right\},~\left\{ M \left\{ M^2 \right\} \right\},~\left\{ M \left\{ M^3 \right\} \right\} \right\}\end{equation}where $\left\{ M \left\{ M^{d} \right\} \right\}$, $d=0,1,2,3$, arerespectively the set of vertices, edges, faces and regions whichdefine the topological entities of the mesh domain. It ispossible to limit the mesh representation to just these entities underthe following restrictions \cite{BeSh97}.\begin{enumerate}\item Regions and faces have no interior holes.\item Each entity of order $d_i$ in a mesh, $M^{d_i}$, may use a particular entity oflower order, $M^{d_j}$, $d_j<d_i$, at most once.\item	For any entity $M^{d_i}_i$ there is a unique set of entities of order $d_i-1$,$\left \{ M^{d_i}_i \left\{M^{d_{i-1}} \right\} \right\}$  that are on the boundary of $M^{d_i}_i$.\end{enumerate}The first restriction means that regions may be directly representedby the faces that bound them, faces may be represented by the edgesthat bound them, and edges may be represented by the vertices thatbound them. The second restriction allows the orientation of an entityto be defined in terms of its boundary entities.For example, the orientation of an edge,$M^1_i$ bounded by vertices $M^0_j$ and $M^0_k$ is uniquely defined asgoing from $M^0_j$ to $M^0_k$ only if $j \neq k$.The third restriction means that a mesh entity is uniquely specifiedby its bounding entities. Most representations including that used inthis paper employ thatrequirement. There are representational schemes where this conditiononly applies to interior entities; entities on the boundary of themodel may have a non-unique set of boundary entities \cite{BeSh97}.A key component of supporting mesh-based simulations is theassociation of the mesh with respect to the geometric model\cite{BeSh97,ShGe92}. This association is referred to asclassification in which the mesh topological entities are classifiedwith respect to the geometric model topological entities upon whichthey lie.{\bf Definition: Classification} - {\it The unique association of meshtopological entities of dimension $d_i$, $M^{d_i}_i$ to thetopological entity of the geometric model of dimension $d_j$,$G^{d_j}_j$ where $d_i \leq d_j$, on which it lies is termedclassification and is denoted $M^{d_i}_i \sqsubseteq G^{d_j}_j$where the classification symbol, $\sqsubseteq$,indicates that the left hand entity, or set, is classified on theright hand entity.}{\bf Definition: Reverse Classification} - {\it For each modelentity, $G^d_j$ , the set of equal order mesh entities classified on thatmodel entity define the reverse classification information for thatmodel entity. Reverse classification is denoted as:}\begin{equation}RC(G^d_j) = \left\{ M^d_i | M^d_i \sqsubseteq G^d_j \right\}\end{equation}The concept of mesh entity classification to a higher levelmodel can be extended to include additional levels of modeldecomposition. Two important cases of this are parallel meshpartitions and structured mesh partitions. In the cases when thesepartitions are non-overlapping the associations are obvious.The concepts can be extended to the case ofoverlapping partitions through the definition of an appropriaterules of the interaction of entities in the different models.Mesh shape information can be effectively associated with thetopological entities defining the mesh. In many cases this is limitedto the coordinates of the mesh vertices and, if they exist, higherorder nodes associated with mesh edges, faces or regions. In addition,it is possible to associate other forms of geometric information withthe mesh entities. For example, the association of Bezier curves andsurface control points with mesh edges and faces for use in p-versionfinite elements \cite{LuSh02}. The mesh classification can be can beused to obtain other needed geometric information such as thecoordinates of a new mesh vertex caused by splitting a mesh edgeclassified on a model face.\subsection{Equation Discretization and the Definition of Solution Fields}The PDEs being solved are written in terms of dependentvariables that are functions of thespace/time domain. For purposes of this discussion, consider the set ofPDEs being solved are written in the form:\begin{equation}{\cal{D}}({\bf u}, \sigma) - f = 0\end{equation}where \begin{itemize}\item $\cal{D}$ represents the appropriate differential operators.\item $\bf{u} (\bf{x},t)$ represents one of more vector dependent variables which arefunctions of the independent variables of space, $\bf{x}$, and time, $t$ .\item $\sigma$ represents one of more scalar dependent variables which arefunctions of the independent variables of space, $\bf{x}$, and time, $t$.\item $f$ represents the forcing functions.\end{itemize}(Note that the complete statement of a PDE problem must include a setof boundary and, for time dependent problems, initial conditions.)In the double discretization process used in mesh-based PDE solvers,the dependent variables are discretized over the individual, orgroups of, mesh entities, either by direct operator discretization(e.g., difference equations) or in terms of a set of basisfunction. In both cases this process specifies a set of distributionfunctions defining how the discretized variables vary over the meshentities and a set of yet to be determined multipliers, called degreesof freedom (dof). The dof can always beassociated with a single mesh entity while the distribution functionsare associated with one or more mesh entities.Three common cases that employ different combinationsof interactions between the mesh entities, the dof and thedistributions are:\begin{itemize}\item Finite difference based on a vertex stencil: In this casethe distribution functions are difference stencilswritten in terms of dof that are the value of the field atspecific neighboring points. The dof are associated with meshvertices. The difference stencil is defined over the mesh entities that link the vertices involved with the stencil.\item Finite volume methods: Finitevolume methods are constructed in terms of distribution functionwritten over individual mesh entities, referred to as cells. In mostcases the field being defined is $C^{-1}$ and the dof are not shared betweenneighboring mesh entities. In this case the dof are associatedwith the mesh entity the distribution is written over. Thecoupling of the dof from different mesh entities is then throughoperators acting over common boundary mesh entities.\item Finite elements with common dof between neighboring elements:Finite element distribution functions, referred to as shape functions,are written over individual mesh entities, referred to as elements. Incases where $C^m,~ m \geq 0$, continuity is required, theshape functions associated with neighboring elements are made $C^m,~m\geq 0$, continuous by having common dof associated with the boundingmesh entities common to the neighboring elements. For example,a $C^0$ field between two neighboring quadraticelements in 2-D can be obtained by using the values of the field at onepoint on the common edge and at the twovertices bounding that edge as dof. In thiscase the full set of dof used by the element distribution function canbe dof associated with any of the mesh entities in the closure of themesh entity of the element. There are other means to meet even higherorder continuity requirements, allof which require sharing dof on the common boundaries.\end{itemize}The process of applying the discretization operation over theappropriate mesh entities will produce a local contribution tothe complete fully discrete system. Theprocesses can be stated symbolically as:\begin{equation}\cal{D} (\bf{D}^c, \bf{d}^c) - \bf{f}^c = 0 \end{equation}where: \begin{itemize}\item $\cal{D}$ represents the discretized differential operators written in termsof appropriate distribution functions, $D^c$, over the domain of thecontributor $C$ and $\bf{d}^c$ represents the vector of dof associated with thatcontributor.\item $\bf{f}^c$ represents the discretized representation of the known``forcing functions'' and boundary conditions for thatcontributor.\end{itemize}The result of the discretization process yields a discreterepresentation of the original PDEs that can be written as:\begin{equation}\bf{k}^c \bf{d}^c = \bf{f}^c\label{eq:contrib_matrix}\end{equation}where $\bf{k}^c$ is a matrix of parameters for contributor $C$ that multiple thevector of dof associated with that contributor, $\bf{d}^c$.The construction of the system contributors can be controlled by theappropriate traversal of information in the high level problemdefinition, or at a level above the mesh such as the mesh patch levelfor structured methods.Note that the solution fields represent the variations of the tensorvariables over the domain of the problem. These fieldsmust be maintained is a form useful for the application of queriesand manipulation as needed for operations that include:\begin{itemize}\item The accurate transfer of the fields to other meshes to provideinput in a multiphysics analysis step, or to maintain the descriptionof the mesh on an adapted field.\item The construction of new fields through operationsthat may project them on new distribution with higher ordercontinuity, combine with other fields, etc.\end{itemize}\subsection{Discretized System Construction and Solution}The relationship of the contributor level discretization given in(\ref{eq:contrib_matrix})to the complete discrete system is dictated by contributor levelmappings that ``map'' the contributor level dof to the``assembled'' vector of the dof for the complete system, $\bf{d}$. Theprocess of constructing the complete system from the contributors isreferred to as the assembly process. Symbolically the completediscrete system can be written as:\begin{equation}\bf{K} \bf{d} = \bf{F}\end{equation}where\begin{itemize}\item $\bf{K}$ is a system level matrix of parameters.\item $\bf{d}$ is the complete vector of dof \item $\bf{F}$ is the complete right hand side vector.\end{itemize}Symbolically the relationship between the contributor level and systemlevel matrices and vectors can be depicted as:\begin{equation}\bf{d} ~=~ A^{N_c}_{c=1}(\bf{d}^c),~K~=~A^{N_c}_{c=1}(\bf{k}^c),~\bf{F}~=~A^{N_c}_{c=1}(\bf{f}^c)\end{equation}where \begin{itemize}\item $N_c$ is the number of contributors in the complete system\item $A^{N_c}_{c=1}$ indicates an assembly operator that is applied to eachcontributors contributions and properly maps it to the completediscrete system.\end{itemize}There are a variety of specific representational forms for thecomplete systems matrices.  The specific form used is function of themethods used to perform the computationally intensive process ofsolving the discrete system to determine the values of the system dof.The global algebraic equations are solved to produce the values of thesystem dofs. Once the system level dof are determined, the mappingsbetween the contributor level and system level dof can be used tocomplete the specification of the solution fields.