\documentclass[10pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}

%opening
\title{Parallel Interface Design for Petascale Parallel Adaptive Simulations}

\author{Ting Xie, Onkar Sahni, Xiaojuan Luo, Min Zhou,   \\ 
        Andrew C. Bauer, Kenneth E. Jansen, Mark Shephard}

\date{\today}

\begin{document}
\maketitle 

\section{Abstract}

Our main goal is to provide the parallel cababilities of the adaptive
control of meshes of billions of elements on petascale
computers.

This paper focuses on entity group and multiple parts per
process, and presents the whole parallel interface design in
SCOREC software framework (FMDB and MeshAdapt), including management of distributed mesh
and field data structures, parallel data
input/output, interface to data partitioning and
load balancing, transfer of data and fields between 
distributed meshes, and predictive load balancing for parallel
adaptive computations. 
All these interfaces are based on distributed memory (MIMD) computing
and message passing (MPI) communication.  


\section{Introduction}     \input{intro.tex}


\section{Partitions}         \input{partitions.tex}


\section{List of Symbols}  \input{symbols.tex}


\section{Mesh Entity Group} \footnote{Only one operation is added in
  this updated document: getBdryEntity.}         


\input{entitygroup.tex}
 

\section{Partition Model}   \input{pmodel.tex}



\section{Using Zoltan to Perform Partition}    \input{zoltan.tex}

   


\section{Dynamic Load Balance and Data Migration}  %\input{lb_migrate.tex}

Static load balance and dynamic load balance, 
Data migration algorithm on multiple parts per process, considering
entity group. 
Attached data migration. 

\section{Parallel File I/O}       %\input{io.tex}

The number of partitioned data files is different from the number of
processes. Multiple partitioned files can be loaded on one process. 

\section{Predictive Load Balance} \input{lb.tex} 

% memory discussion. 


\section{Conclusion}


\begin{thebibliography}{9}
\bibitem{LLNL} Steven F.Ashby. \textit{Scalable Algorithms and Novel
  Paradigms for Petascale Simulation}. Feb, 2006.
\bibitem{TACC} H. Carter Edwards. \textit{TACC Seminar:
SIERRA Framework Supporting Parallel Adaptive Multiphysics
Applications}. May, 2004. 
\bibitem{devine} Karen Devine. \textit{Partitioning and Dynamic Load
  Balancing for Petascale Applications}. SciDAC, 2007.
\bibitem{ITAPS} ITAPS. \textit{ITAPS Parallel Interface: Requirements
  Document}. Jun 12, 2007. 
\bibitem{Seol} Seegyoung Seol. \textit{FMDB:Flexible Distributed Mesh
  Database for Parallel Automated Adaptive Analysis}. PhD dissertation
  in RPI, 2005.
\bibitem{Zoltan} \textit{Zoltan: Parallel Partitioning, Load Balancing and Data-Management
Services}. http://www.cs.sandia.gov/Zoltan/.
\bibitem{R} Loy R. \textit{Autopack User Mannual}. Science Division
  Argonne National Laboratory, 2000. 
\bibitem{Pain} C.C.Pain, A.P.Umpleby, C.R.E.de Oliveira and
  etc. \textit{Tetrahedron mesh optimization and adaptivity for steady-state and transient finite
element calculations}. Computer methods in applied mechanics and
engineering. 2001.
\bibitem{Li} X.Li, \textit{Mesh Modification procedures for general 3-D non-manifold domains}. PhD thesis, Rensselaer Polytechnic Institute, August 2003.
\bibitem{Flaherty} J.E.Flaherty, R.M.Loy, P.C.Scully, M.S.Shephard and
  etc. \textit{Load Balancing and communication Optimization for
  Parallel Adaptive Finite Element Methods}, SCCC'97.

\end{thebibliography}




\end{document}


