\documentclass{aiaa-tc}
\usepackage{pslatex}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\pagestyle{empty}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{float}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage{multicol}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users
\providecommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\renewcommand{\floatpagefraction}{1}
\renewcommand{\textfraction}{0}
\renewcommand{\topfraction}{0.9}
\setcounter{topnumber}{4}
\usepackage{url}

\newenvironment{lyxlist}[1]
 {\begin{list}{}
   {\settowidth{\labelwidth}{#1}
    \setlength{\leftmargin}{\labelwidth}
    \addtolength{\leftmargin}{\labelsep}
    \renewcommand{\makelabel}[1]{##1\hfil}}}
 {\end{list}}
 \newenvironment{lyxcode}
   {\begin{list}{}{
     \setlength{\rightmargin}{\leftmargin}
     \setlength{\listparindent}{0pt}% needed for AMS classes
     \raggedright
     \setlength{\itemsep}{0pt}
     \setlength{\parsep}{0pt}
     \normalfont\ttfamily}%
    \item[]}
   {\end{list}}

\begin{document}

\title{An Interoperable Mesh Component for PDE Simulation Development}


\author{Lori Freitag Diachin\thanks{Center for Applied Scientific
  Computing, Lawrence Livermore National Laboratory} \and Carl
  Ollivier-Gooch\thanks{Advanced Numerical Simulation Laboratory,
  University of British Columbia} \and Mark
  S. Shephard\thanks{Scientific Computation Research Center, Rensselaer
  Polytechnic Institute } \and Timothy Tautges\thanksibid{Mathematics
  and Computer Science Division, Argonne National Laboratory} \and
  Harold Trease\thanks{Pacific Northwest National Laboratory} }

\maketitle
\begin{abstract}

PDE-based numerical simulation applications commonly use basic software
infrastructure to manage mesh, geometry, and discretization data. The
commonality of this infrastructure implies the software is theoretically
amenable to re-use. However, the traditional reliance on library-based
implementations of these functionalities hampers experimentation with
different software instances that provide similar functionality.  This
is especially true for meshing and geometry libraries where applications
often directly access the underlying data structures, which can be quite
different from implementation to implementation.  Thus, using different
libraries interchangeably or interoperably for this functionality has
proven difficult at best and has hampered the wide spread use of
advanced meshing and geometry tools developed by the research community.
To address these issues, the Terascale Simulation Tools and Technologies
center is working to develop standard interfaces to enable the creation
of interoperable and interchangeable simulation tools. In this paper, we
focus on a language- and data-structure-independent interface supporting
query and modification of mesh data conforming to a general abstract
data model. We describe the model and interface, and provide programming
``best practices'' recommendations based on early experience
implementing and using the interface.

\end{abstract}

\input{intro.tex}

\input{design.tex}

\input{interface-desc.tex}

\input{best-practices.tex}

\input{conclusions.tex}

\section*{Acknowledgements}

This work was performed under the auspices of the U.S. Department of
Energy by the University of California Lawrence Livermore National
Laboratory under contract No. W-7405-Eng-48 (UCRL-JRNL-213577); by The
University of British Columbia under Canadian Natural Sciences and
Engineering Research Council (NSERC) Special Research Opportunities
Grant SRO-299160; and by Rensselaer Polytechnic Institute under DOE
grant number DE-FC02-01ER25460.

\bibliographystyle{plain}
\bibliography{../biblio/tstt}

\input{appendix.tex}

\end{document}
